StarCraft 2 Replay Data Exploration by Benjamin Xiao
========================================================

> **Tip**: You will see quoted sections like this throughout the template to
help you construct your report. Make sure that you remove these notes before
you finish and submit your project!

> **Tip**: One of the requirements of this project is that your code follows
good formatting techniques, including limiting your lines to 80 characters or
less. If you're using RStudio, go into Preferences \> Code \> Display to set up
a margin line to help you keep track of this guideline!

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(dplyr)
library(knitr)
library(gridExtra)
```

```{r}
# for tiny
setwd("~/dand_files/project04_files/")

# for mini 
setwd("~/dand_files/project04/")
```


```{r echo=FALSE, Load_the_Data}
# Load the Data
sc <- read.csv("starcraft.csv")
```

# Introduction

StarCraft 2 is a sequel to the ever popular real time strategy game, StarCraft Brood War. In its competitive form, two players face each other on various maps choosing one of three available races or a random choice. The database used can be downloaded here: https://www.kaggle.com/sfu-summit/starcraft-ii-replay-analysis and the original study of this dataset is at http://summit.sfu.ca/item/13328

I am exploring this dataset to look mehanical differences among players of different skill levels. The skill levels are broadly placed into bins from 1 to 8, 8 being the most skilled and smallest percentage of the sample. 

Each variable in the dataset is hypothesized to be correlated with higher skill. For instance, better players will make and be able to use more complex units. Lower level players generally play with lower APM and utilize less mechanical tools available (hot keys and mini map functions). Most variables are divided by the time elapsed to generate a standard for variables that  naturally acculumate over time. 

This data is also a snapshot of the overall skill levels for players at the time of data collection and for the specific expansion called Wings of Liberty. The explanation of league placement is shown here: http://wiki.teamliquid.net/starcraft2/Battle.net_Leagues

My main question is which variables affect skill level most? Other questions include what habits separates players in top placements? This can provide insight into where players at each skill level hit skill walls and must work deliberately in order to improve and move into higher league placements.

# Univariate Plots Section

> **Tip**: In this section, you should perform some preliminary exploration of
your dataset. Run some summaries of the data and create univariate plots to
understand the structure of the individual variables in your dataset. Don't
forget to add a comment after each plot or closely-related group of plots!
There should be multiple code chunks and text sections; the first one below is
just to help you get started.

```{r echo=FALSE, Univariate_Plots}
summary(sc)

ggplot(aes(LeagueIndex), data = sc) + 
  geom_histogram()
table(sc$LeagueIndex)

ggplot(aes(Age), data = sc) + 
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = seq(0,50,5))
summary(sc$Age)
table(sc$Age)
```
These are counts of the contributing players to this dataset. Technically, professional and grandmaster players are both in the grandmaster category, but the community agrees pro players are in a different class.

```{r Time_data}
ggplot(aes(HoursPerWeek), data = sc) + 
  geom_histogram()

ggplot(aes(HoursPerWeek), data = sc) + 
  geom_histogram(binwidth = 1)

ggplot(aes(HoursPerWeek), data = sc) + 
  geom_histogram(binwidth = 1) + 
  scale_x_continuous(limits = c(0, 50), breaks = seq(0, 50, 5))

summary(sc$HoursPerWeek)

ggplot(aes(HoursPerWeek), 
       data = subset(sc, sc$HoursPerWeek < 112)) + 
  geom_histogram(binwidth = 1)

summary(subset(sc$HoursPerWeek, sc$HoursPerWeek < 112))

ggplot(aes(TotalHours), data = sc) + 
  geom_histogram()

summary(sc$TotalHours)

ggplot(aes(TotalHours), data = sc) + 
  geom_histogram(binwidth = 50) + 
  xlim(0, 1500)

summary(subset(sc$TotalHours, sc$TotalHours < 15000))
```
The hours spent playing StarCraft 2 is left skewed with many outliers. Zooming into hour per week, we can see most players play between 0 to 35 hours a week. Zooming into total hours, most players are in the 0 to 1500 hours categories. The higher end of the total hours played is just impossible, so I subsetted the data in the final summary. 

```{r Univariate_Plots_APM}
ggplot(aes(APM), data = sc) +
  geom_histogram()

ggplot(aes(APM), data = sc) + 
  geom_histogram(binwidth = 5)

ggplot(aes(APM), data = sc) + 
  geom_histogram(binwidth = 5) + 
  scale_x_continuous(limits = c(30, 220))

summary(sc$APM)
```
The majority of APM values are somewhere between 50 and 200. Zooming in, we get a clearer view of that most players do play within the 80 and 150 APM range. 

```{r Univariate_Plots_general}
ggplot(aes(GapBetweenPACs), data = sc) + 
  geom_histogram(binwidth = 1)

ggplot(aes(GapBetweenPACs), data = sc) + 
  geom_histogram(binwidth = 1) + 
  scale_x_continuous(limits = c(0, 150))

summary(sc$GapBetweenPACs)

ggplot(aes(TotalMapExplored), data = sc) + 
  geom_histogram(binwidth = .00001)

summary(sc$TotalMapExplored)

ggplot(aes(WorkersMade), data = sc) + 
  geom_histogram(binwidth = 0.0001)

summary(sc$WorkersMade)
```
These are general, singular measurements of different skill levels. Of course, these are all naturally correlated with APM, so it makes sense that the distributions are similar to the APM histogram. Since these are all standardized based on a time period, it also makes sense that they have similar distributions based on this normalization. 

```{r Workers_created_raw}
sc$WorkersCreatedTotal <- sc$WorkersMade * sc$MaxTimeStamp
sc$WorkersCreatedTotal <- round(sc$WorkersCreated)
summary(sc$WorkersCreatedTotal)
```


> **Tip**: Make sure that you leave a blank line between the start / end of
each code block and the end / start of your Markdown text so that it is
formatted nicely in the knitted text. Note as well that text on consecutive
lines is treated as a single space. Make sure you have a blank line between
your paragraphs so that they too are formatted for easy readability.

# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?

There are 3395 contributing players to this dataset with one general skill categorization (LeagueIndex), player information (Age, HoursPerWeek, TotalHours), and general measurements of skill. 

LeagueIndex goes from 1 to 8. 1 being the worst and 8 being you should go compete in tournaments.

Observations:
* Most contributing users are in the top half of the player pool according to the league placement distribution.
* Mean APM among these players is 117.05 which is a little low from my expectations
* All the variables that have been divided by time are very similarly distributed.
* The max value in TotalHours is 1000000 hours which is definitely an error because that is physically impossible. The game was released in July 2010, and the data was collected in 2013. This is probably just some people being funny.

### What is/are the main feature(s) of interest in your dataset?

I'd like to primarily see which measure or collection of measurements characterize skill best. I suspect all PAC data is the most telling and GapBetweenPACs to be the difference before players achieve GrandMaster or higher placement. Secondarily, I'd like to explore the relationship with time spent playing to help debunk the idea that ability in StarCraft 2 is fixed. 

### What other features in the dataset do you think will help support your \ 
investigation into your feature(s) of interest?

I'm guessing hot key usage is the next best predictor of skill. Habitually using your hot keys allows a player to play the game more fluidly, faster, and manage various aspects of the game simultaenously. Neglecting hot keys, particularly the SelectByHotkeys significantly hinders overall gameplay.

### Did you create any new variables from existing variables in the dataset?

I created a total workers made variable by multiplying WorkersMade and MaxTimeStamp. Since there is no such thing as a fraction of a worker, I used the round() function to make everything integers instead.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

I subsetted TotalHours and HoursPerWeek variables to eliminate the impossible values such as playing for 100000 when the game has not been released for that long. I subsetted HoursPerWeek to values less than 112 hours because there are only 168 hours in a week, and I'm assuming a third of those hours are spent sleeping or doing other things besides playing StarCraft 2. 

# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!
>>>>>>> 4248a1537320273e9c25338bbff21d74d1323c48

Something to improve on - APM patterns through the game because peak APM may be more telling than regular APM through the game if all you do is spam.
